---
title: "source_summaries"
author: "Chris Roeder"
date: "6/11/2019"
output:
  html_document: default
  pdf_document: default
---

This notebook summarizes data from CSV files from either end of the harmonization process to help verify that process. There is a major section describing the input from each of four studies: ACCORD, AIM-HIGH, ALLHAT and BARI-2D. Finally there is a section describing the output CSV that has a column for each study. There are special sections to show calculation resutls for antihyp and dthDays.
The output is written to create a knitted (html or pdf) document. The formatting doesnt look as good in RStudio

```{r setup, include=TRUE, cache=FALSE}
require("knitr")
opts_knit$set(root.dir="/mnt/workspace/test/studies")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment=NA)
#knitr::opts_chunk$set(results='asis') #   R code that outputs HTML (table1) needs NOT to have its output sourrounded with <pre> etc.
#install.packages("table1")
library(table1)
```

# ACCORD
```{r ACCORD, results="asis"}

# The following lists have a 1:1 correspondance so the 3rd vector in the columns list contains colums for the third file in the files vector.
# Facs lists columns that should be converted to factors. Where_clause indicates which files should have an SQL "where" clause applied.
files <- c(
    'ACCORD/3-Data Sets - Analysis/3a-Analysis Data Sets/ACCORD_KEY.csv',
    'ACCORD/3-Data Sets - Analysis/3a-Analysis Data Sets/OTHERLABS_201604.csv',
    'ACCORD/3-Data Sets - Analysis/3a-Analysis Data Sets/BLOODPRESSURE_201604.csv',
    'ACCORD/4-Data Sets - CRFs/4a-CRF Data Sets/F01_INCLUSIONEXCLUSIONSUMMARY.csv',
    'ACCORD/4-Data Sets - CRFs/4a-CRF Data Sets/F07_BASELINEHISTORYPHYSICALEXAM.csv',
    'ACCORD/3-Data Sets - Analysis/3a-Analysis Data Sets/LIPIDS_201604.csv',
    'ACCORD/3-Data Sets - Analysis/3a-Analysis Data Sets/CONCOMITANTMEDS.csv',
    'ACCORD/3-Data Sets - Analysis/3a-Analysis Data Sets/CVDOUTCOMES_201604.csv')

columns <- list(
    c('raceclass', 'MaskID', 'baseline_age', 'female'),
    c('fpg'), # where
    c('hr', 'dbp', 'sbp'), # where
    c('x2mi', 'x2stroke', 'cabg', 'ptci'),
    c('wt_kg', 'ht_cm', 'cigarett', 'yrstens'),
    c('ldl', 'hdl', 'chol', 'trig'), # where
    c('statin', 'aspirin', 'loop', 'thiazide', 'ksparing', 'a2rb', 'acei', 'dhp_ccb', 'nondhp_ccb', 'beta_blocker','reserpine', 'other_bpmed', 'nitrate','vasodilator','central_agent', 'alpha_blocker', 'potassium', 'other_diabmed','other_lipidmed'), # where
    c('fuyrs_tm', 'censor_tm')
)
facs <- list(
    c('raceclass', 'female'), # KEY
    c(), # OTHERLABS
    c(), # BLOODPRESSURE
    c('x2mi', 'x2stroke','cabg','ptci'), # F01
    c('cigarett'), # F07
    c(), #LIPIDS
    c('statin', 'aspirin', 'loop', 'thiazide', 'ksparing', 'a2rb', 'acei', 'dhp_ccb', 'nondhp_ccb', 'beta_blocker','reserpine', 'other_bpmed', 'nitrate','vasodilator','central_agent', 'alpha_blocker', 'potassium', 'other_diabmed','other_lipidmed'), #CON
    c('censor_tm') #CVDOUTCOMES
)
where_clause <- list(F, T, T, F, T, T, T, F)


i=1
for (file in files) {
    cat("<br><b>"); cat(file); cat("</b><br>")
    file_data <- read.csv(file)
    if (where_clause[[i]]) {
        file_data <- file_data[file_data$Visit=='BLR',]
    }
    file_data_no_factor <- file_data
    for (fac_name in facs[[i]]){
        if (!is.null(fac_name)) {
            file_data[,fac_name]=factor(file_data[,fac_name])
        }
    }
    

    if (file ==  'ACCORD/3-Data Sets - Analysis/3a-Analysis Data Sets/CVDOUTCOMES_201604.csv') {
        # convert years to days for death days here
        file_data$fuyrs_tm = file_data$fuyrs_tm * 365.25
        
        # do a table1 for death/censor times 
        cat("<br><b>Death/Censor table</b><br>")
        death_fmla = '~fuyrs_tm |  censor_tm'
        cat(table1(as.formula(death_fmla), data=file_data, output="markdown"))
        cat("<br>")
        cat("<br><b>"); cat(file); cat("</b><br>")
    }
    
    
    fmla <- paste('~', paste(columns[[i]], collapse=" + "))
    cat(table1(as.formula(fmla), data=file_data, output="markdown"))
    cat("<br>")

    # antihyp
    if (file == 'ACCORD/3-Data Sets - Analysis/3a-Analysis Data Sets/CONCOMITANTMEDS.csv') {
        file_data_no_factor$antihyp = 0

        # replace NA with FALSE
        ah_cols = c("nitrate", "reserpine", "thiazide", "beta_blocker", "acei", "a2rb", "alpha_blocker", "central_agent", "loop", "dhp_ccb", "nondhp_ccb", "ksparing", "other_bpmed")

        file_data_no_factor[ah_cols][is.na(file_data_no_factor[ah_cols])] <- FALSE
        file_data_no_factor$antihyp = (
      file_data$vasodilator  == 1
    | file_data_no_factor$nitrate  == 1
    | file_data_no_factor$reserpine  == 1
    | file_data_no_factor$thiazide  == 1
    | file_data_no_factor$beta_blocker  == 1
    | file_data_no_factor$acei  == 1
    | file_data_no_factor$a2rb  == 1
    | file_data_no_factor$alpha_blocker == 1
    | file_data_no_factor$central_agent  == 1
    | file_data_no_factor$loop  == 1
    | file_data_no_factor$dhp_ccb  == 1
    | file_data_no_factor$ksparing  == 1
    | file_data_no_factor$nondhp_ccb  == 1
    | file_data_no_factor$other_bpmed  == 1    )
        cat("<br><br><b>ACCORD antihyp:</b><br>")
        cat(sum(as.numeric(file_data_no_factor$antihyp), na.rm=TRUE))
        cat("<br><br>")
    }
    i=i+1
}
```

# AIM-HIGH
```{r AIMHIGH,  results='asis'}

# The following lists have a 1:1 correspondance so the 3rd vector in the columns list contains colums for the third file in the files vector.
# Facs lists columns that should be converted to factors. Where_clause indicates which files should have an SQL "where" clause applied.

files <- c(
    'AIM-HIGH/SASData/Tabulations/PE.csv',
    'AIM-HIGH/SASData/Tabulations/HX.csv',
    'AIM-HIGH/SASData/Tabulations/FOLLA6.csv',
    'AIM-HIGH/SASData/Tabulations/EXC.csv',
    'AIM-HIGH/SASData/Tabulations/DEMO.csv',
    'AIM-HIGH/SASData/Analysis/LAB.csv',
    'AIM-HIGH/SASData/Analysis/PATIENT.csv',
    'AIM-HIGH/SASData/Tabulations/CONMED.csv',
    'AIM-HIGH/SASData/Tabulations/INCL.csv',
    'AIM-HIGH/SASData/Tabulations/DEATH.csv'
   )

columns <- list(
    c('PE_Wt', 'PE_Ht','PE_BPsys', 'PE_BPdia'),
    c('Hx_Hyper', 'Hx_MI', 'Hx_Stroke', 'Hx_CABG', 'Hx_PCI'),
    c(),
    
    c(),
    c(),
    c('Result'), #lab test=CHOL visit=0
    
    c('Age', 'Sex', 'Race', 'BMI', 'HxDiab', 'SysBP', 'DiasBP', 'LDL_C_Base', 'HDL_C_Base', 'Trig_Base', 'DtCensor', 'PatientKey', 'Smoke'),
    c('CM_Aspirin', 'CM_ATII', 'CM_Ace', 'CM_Beta','CM_Nit','CM_Diu', 'CM_CCB', 'CM_Metformin', 'CM_Plavix','CM_Insulin'), # visit=Baseline
    c('Incl_CVD'),
    
    c('DTH_Date')
)
facs <- list(
    c(),
    c('Hx_Hyper', 'Hx_MI', 'Hx_Stroke', 'Hx_CABG', 'Hx_PCI'),
    c(),
    c(),
    c(),
    c(), #lab test=CHOL visit=0
    c('Sex', 'Race', 'HxDiab',  'Smoke'),
    c('CM_Aspirin', 'CM_ATII', 'CM_Ace', 'CM_Beta','CM_Nit','CM_Diu', 'CM_CCB', 'CM_Metformin', 'CM_Plavix','CM_Insulin'), # visit=Baseline
    c('Incl_CVD'),
    c('DTH_Date')
)
where_clause <- list('Visit', 'no',  'no','no','no',  'Visit', 'no', 'visit', 'no', 'no')


i=1
for (file in files) {
    cat("<br><b>"); cat(file); cat("</b><br>")
    file_data <- read.csv(file)
    print(names(file_data))
    if (length(columns[[i]]) > 0){

        # Where Clause
        if (where_clause[[i]] != 'no') {
           if (where_clause[[i]] =='Visit') {
            file_data <- file_data[file_data$Visit=='Baseline',]
           } else {
            file_data <- file_data[file_data$visit=='Baseline',]
           }
        }

        # save unfactored version of data for use below
        data <- file_data

        # factor conversions
        for (fac_name in facs[[i]]){
            if (!is.null(fac_name)) {
                a=factor(file_data[[fac_name]])
                file_data[[fac_name]]=a
            }
        }
        fmla <- paste('~', paste(columns[[i]], collapse=" + "))
        cat(table1(as.formula(fmla), data=file_data, output="markdown"))
    } else {
        cat("no columns")
    }
    cat("<br>")

    if (file == 'AIM-HIGH/SASData/Tabulations/CONMED.csv') {
        # remove NA
        cols <- c('CM_Nit', 'CM_Diu', 'CM_Beta', 'CM_Ace', 'CM_ATII', 'CM_CCB')
        data[cols][is.na(data[cols])] <- 0

        data$antihyp = 0
        data$antihyp = (
              data[,'CM_Nit'] == 1
            | data[,'CM_Diu'] == 1
            | data[,'CM_Beta'] == 1
            | data[,'CM_Ace'] == 1
            | data[,'CM_ATII'] == 1
            | data[,'CM_CCB'] == 1)
        cat("<br><br><b>AIM-HIGH antihyp</b><br>")
        cat(sum(as.numeric(data$antihyp), na.rm=TRUE))
        cat("<br><br>")
    }
    if (file == 'AIM-HIGH/SASData/Tabulations/DEATH.csv') {
        death_data = data
    }
    if (file == 'AIM-HIGH/SASData/Analysis/PATIENT.csv') {
        patient_data = data
    }

    i=i+1
}

# do a table1 for death/censor times 
joined_data = merge(x=patient_data, y=death_data, by='PatientKey', all=T)
cols=c("DTH_Date")
joined_data[cols][!is.na(joined_data[cols])] <- 1
joined_data[cols][is.na(joined_data[cols])] <- 0

 
joined_data[['DTH_Date']]=factor(joined_data[['DTH_Date']])
cat("<br>")
    cat(unique(joined_data$DTH_Date))
cat("<br><b>Death/Censor table</b><br>")
death_fmla = '~DtCensor |  DTH_Date'
cat(table1(as.formula(death_fmla), data=joined_data, output="markdown"))
cat("<br>")
cat("<br><b>"); cat(file); cat("</b><br>")
    


```

# ALLHAT
```{r ALLHAT,  results='asis'}

# The following lists have a 1:1 correspondance so the 3rd vector in the columns list contains colums for the third file in the files vector.
# Facs lists columns that should be converted to factors. Where_clause indicates which files should have an SQL "where" clause applied.

files <- c(
       'ALLHAT/DATA/Summary/DOX_VSC.csv',
       'ALLHAT/DATA/Summary/LLT_VSC.csv'
   )

columns <- list(
    c('llt',
    'blbmi',
    'blhgt',
    'blwgt',
    'sex',
    'hispanic',
    'aspirin',
    'dydeath',
    'age',
    'mistroke',
    'blmeds2',
    'dbp6m12',
    'bv2sbp',
    'blgfr',
    'achol',
    'bv2dbp',
    'rzgroup',
    'cursmoke',
    'sbp6m12',
    'hxcabg',
    'ahdl',
    'aldl',
    'aftrig',
    'race',
    'race2',
  #  'index',
    'death',
    'diabetes'),
  c('lrzgrp')
)
facs <- list(
    c('llt',
    'sex',
    'hispanic',
    'aspirin',
    'mistroke',
    'blmeds2',
    'rzgroup',
    'cursmoke',
    'hxcabg',
    'race',
    'race2',
  #  'index',
    'death',
    'diabetes'),
  c('lrzgrp')
)


i=1
for (file in files) {
    cat("<br><b>"); cat(file); cat("</b><br>")
    file_data <- read.csv(file)
    if (length(columns[[i]]) > 0){
        for (fac_name in facs[[i]]){
            if (!is.null(fac_name)) {
                file_data[,toupper(fac_name)]=factor(file_data[,toupper(fac_name)])
            }
        }

        fmla <- paste('~', paste(toupper(columns[[i]]), collapse=" + "))
        cat(table1(as.formula(fmla), data=file_data, output="markdown"))
    }

    if (i==1) { # only for the DOX_VSC file
        cat("<br><b>ALLHAT antihyp (blmeds2)</b><br>")
        cat(sum(as.numeric(file_data$BLMEDS2) -2, na.rm=TRUE) * -1) # values are 1 and 2 for yes and no,
        cat("<br>")
        
        # do a table1 for death/censor times 
        cat("<br><b>Death/Censor table</b><br>")
        death_fmla = '~DYDEATH |  DEATH'
        cat(table1(as.formula(death_fmla), data=file_data, output="markdown"))
        cat("<br>")
        cat("<br><b>"); cat(file); cat("</b><br>")
    }
    i=i+1
}
```

# BARI-2D
```{r BARI2D,  results='asis'}
files <- c('BARI-2D/data/BARI2D_BL.csv',
           'BARI-2D/data/BARI2D_ENDPTS.csv',
           'BARI-2D/data/BARI2D_LONG.csv'
           #, 'BARI-2D/data/BARI2D_PE.csv'
           )

# The following lists have a 1:1 correspondance so the 3rd vector in the columns list contains colums for the third file in the files vector.
# Facs lists columns that should be converted to factors. Where_clause indicates which files should have an SQL "where" clause applied.

columns <- list(
  c('smkcat',  #BL
     'bmi',    #BL
     'priorrev',#BL
     'ccb',
     'hxhtn',  #BL
     'asp',    #BL
     'race',   #BL
 #    'id',
     'dbp_sit',#BL
     'pulse60_sit',
     'sbp_sit',#BL
     'height', #BL
     'diur',
     'acei',   #BL
     'arb',    #BL tzd    #BL
     'nsgn',   #BL
     'hba1c',
     'hdl',    #BL
     'ldl',    #BL
     'trig',   #BL
     'tchol',  #BL
     'sex',    #BL
     'weight', #BL
     'age',    #BL
     'betab',  #BL
     'hmg',    #BL
     'hxmi',   #BL
     'hxcva'), #BL
  c('deathfu','death'), #ENDPOINTS
  c('omega',   #
    'fibr',
    'niac',
    'sbp_sit',
    'dpp',
    'amylin',
    'dbp_sit',
    'insulin',
    'apa',
    'sulf',
    'biguanide',
    'incrmim')
)
facs <- list(
  c('smkcat',  #BL
     'priorrev',#BL
     'ccb',
     'hxhtn',  #BL
     'asp',    #BL
     'race',   #BL
     'diur',
     'acei',   #BL
     'arb',    #BL
     'nsgn',   #BL
     'sex',    #BL
     'betab',  #BL
     'hmg',    #BL
     'hxmi',   #BL
     'hxcva'), #BL
  c('death'), #ENDPOINTS
  c('omega',
    'fibr',
    'niac',
    'dpp',
    'amylin',
    'insulin',
    'apa',
    'sulf',
    'biguanide',
    'incrmim')
)

nums <- list(
    c('height','weight'), c(), c('sbp_sit', 'dbp_sit')
)


i=1
for (file in files) {

    cat("<br><b>"); cat(file); cat("</b><br>")
    if (!file.exists(file)) {
        cat(paste("no good ", file))
    } else {
        file_data <- read.csv(file)

        # filter by fuper=-1 for BARI2D_LONG
        if (file == 'BARI-2D/data/BARI2D_LONG.csv') {
                        file_data <- file_data[file_data$fuper==-1,]
        }

        # do antihyp when in BARI2D_BL
        if (file == 'BARI-2D/data/BARI2D_BL.csv') {
            # remove NA
            cols <- c('diur', 'betab', 'acei', 'arb', 'ccb', 'nsgn')
            file_data[cols][is.na(file_data[cols])] <- '0'
            # antihyp
            file_data$antihyp=0
            file_data$antihyp = (
                  file_data[,'diur']=='1'
                | file_data[,'betab']=='1'
                | file_data[,'acei']=='1'
                | file_data[,'arb']=='1'
                | file_data[,'ccb']=='1'
                | file_data[,'nsgn']=='1')
            cat("<br><b>BARI-2D antihyp</b><br>")
            cat(sum(file_data$antihyp)) #, na.rm=TRUE))
            cat("<br>")
            cat(summary(file_data$antihyp))
        }
        
        if (file == 'BARI-2D/data/BARI2D_ENDPTS.csv') {
            cat("<br><b>Death/Censor table</b><br>")
            death_fmla = '~deathfu |  death'
            file_data[['death']]=factor(file_data[['death']])
            cat(table1(as.formula(death_fmla), data=file_data, output="markdown"))
            cat("<br>")
            cat("<br><b>"); cat(file); cat("</b><br>")
        }
        
        if (length(columns[[i]]) > 0){
            for (fac_name in facs[[i]]){
                if (!is.null(fac_name)) {
                    file_data[,fac_name]=factor(file_data[,fac_name])
                }
            }
            for (col_name in nums[[i]]) {
                if (!is.null(col_name)) {
                    file_data[,col_name]=as.numeric(file_data[,col_name])
                }
            }
        }


        fmla <- paste('~', paste(columns[[i]], collapse=" + "))
        cat(table1(as.formula(fmla), data=file_data, output="markdown"))
    }
    i=i+1
}



```

# COMBINED, HARMONIZED
```{r OMOP,  results='asis'}
files=c('/mnt/workspace/output/accord.csv', '/mnt/workspace/output/aimhigh.csv', '/mnt/workspace/output/allhat.csv', '/mnt/workspace/output/bari2d.csv')
df_all <- data.frame()
for (file in files) {
    cat("<br><b>"); cat(file); cat("</b><br>")
    file_data <- read.csv(file)
    df_all <- rbind(df_all, file_data)
}

col_names <- lapply(names(df_all), function(y) Filter(function(x) x!='studyIDI' && x!='study.1' && x!='character(0)', y))
cols_length = length(col_names)
col_names <- c(col_names[1:46],col_names[48], col_names[49])
fmla <- paste('~', paste(col_names, collapse=" + "))
fmla <- paste(fmla, '| study.1', sep='')

factor_cols = c('Sex', 'Race', 'Race2', 'Toba', 'Htn', 'HxDM', 'HxMIStr', 'revasc', 'HMG', 'asprin', 'antihyp', 
    'MI', 'HxStr', 'CABG', 'PCI', 'Loop', 'thiazide', 'KDiur', 'ARB', 'AceI', 'Dihy', 'nonDihy', 'BetaB',
    'rese', 'othBPmed', 'ONitra', 'nslNitrate', 'VASO', 'Cent', 'AlphaB', 'K', 'dthStat', 'study', 'DIUR',
    'CaCB', 'TZD')
for (col in factor_cols) {
    df_all[,col] = factor(df_all[,col])
}

# calculate dthDays outside of Table1....because they are different
cat("<br><b>dthDays</b><br>")
study_names = unique(df_all$study.1)
for (s_name in study_names) {
s_data <- df_all[df_all$study.1==s_name,]
    cat(paste("<b>", s_name, "</b>"))
    cat(paste(' mean', mean(s_data[,'dthDays'], na.rm=TRUE)))
    cat(paste(' median', median(s_data[,'dthDays'], na.rm=TRUE)))
    cat(paste(' min', min(s_data[,'dthDays'], na.rm=TRUE)))
    cat(paste(' max', max(s_data[,'dthDays'], na.rm=TRUE)))
    cat("<br>")
}

cat("<br><b>Death/Censor table</b><br>")
cols <- c('dthStat')
df_all[cols][is.na(df_all[cols])] <- '0'
death_fmla = '~dthDays | study.1 + dthStat'
cat(table1(as.formula(death_fmla), data=df_all, output="markdown"))


cat(table1(as.formula(fmla), data=df_all, output="markdown"))

# split dthDays between dthStatus values
```

